---
title: "VPS_ValidationReportTempalte"
format: 
  html:
    fig-align: center
    fig-width: 4
    fig-height: 4
    fig-format: png
execute:
  echo: true
  cache: false

tidy: TRUE
toc: true
toc-location: left
number-sections: true
number-depth: 3

    
params:
  name: 
    label: "Put your name here"
    value: "SuperUser"
    input: text
  experiment: 
    label: "Experiment name (used as report title)"
    value: "Calibration curve 1"
    input: text
  experiment_date:
    label: "Date of experiment"
    #format: "dd-mm-yyyy"
    value: "`r Sys.Date()`"
    input: date
  comp:
    label: "Compound"
    value: "amoxicillin"
    input: text  
  conc_units:
    label: "Concentration units"
    value: "ppm"
  #IS: 
   # label: "Internal standard"
   # value: "amoxicillin D4"
  #  input: text
 # IS_conc:
 #   label: "Concentration of internal standard"
 #   value: 1 
 #   input: text

  matrix:
    label: "Sample matrix"
    value: "plasma"
    input: text
  operators: 
    label: "Operators"
    value: "Pieter-Jan De Sutter"
    input: text 
  study: 
    label: "Study"
    value: "LMBKA project"
    input: text
  file1:
    label: "Select Excell file (USE TEMPLATE)"
    value: "Test_input_06jan25.xlsx"
    input: file
    placeholder: "Please select a file"
  #weighing:
    #label: "Select weighing function"
    #value: 2
    #input: radio
   # choiceNames: [No weighing function, 1/x, 1/x²] 
   # choiceValues: [1,2,3]
  StdReplicates:
    label: "Select Number of Standard Replicates to Process "
    value: 2
    input: number
    min: 1
    max: 4
    step: 1
#  LogC:
 #   label : "Use log scale for concentration and response plots"
 #   value: TRUE
  decimalsC:
    label: "Decimals for concentration output"
    value: 2
    input: slider
    min: 0
    max: 4
    step: 1
  decimalsR:
    label: "Decimals for response output"
    value: 2
    input: slider
    min: 0
    max: 6
    step: 1
  SizeFont:
    label: "Font size tables"
    input: slider
    value: 14
    min: 8
    max: 16
    step: 1
  SizeFig:
    label: "Font size graphs"
    input: slider
    value: 12
    min: 8
    max: 16
    step: 1
---

```{r}
#| label: load library

library(tidyverse)
library(readxl)
library(kableExtra)
library(gridExtra)
library(broom)
library(car)
library(plotly)
library(ggplot2)
library(viridis)
library(ggsci)
library(gt)
```

```{r}
#| label: Set params as objects
# function to calculate %RSD
rsd <- function(x,
                as.percentage = TRUE,
                na.rm = TRUE,
                adjust = FALSE,
                adjust.lo = 1) {

  if (adjust) {
    if (any(x < adjust.lo)) {
      x <- x - min(x, na.rm = TRUE)
    }
  }

  cov <- sd(x, na.rm = na.rm) / mean(x, na.rm = na.rm)
  if (as.percentage) cov <- cov * 100
  cov
}

 #custom function to transpose while preserving names
transpose_df <- function(df) {
  t_df <- data.table::transpose(df)
  colnames(t_df) <- rownames(df)
  rownames(t_df) <- colnames(df)
  t_df <- t_df %>%
    tibble::rownames_to_column(.data = .) %>%
    tibble::as_tibble(.)
    return(t_df)}


decimalsC <- params$decimalsC
decimalsR <- params$decimalsR
Std_Replicates <- params$StdReplicates
font_size <- params$SizeFont
plot_size <- params$SizeFig
```

```{r}
#|label: read data file
DATA <- read_excel(params$file1)

#Make a new column that combines type, level, replicate
DATA <- unite(DATA,c(Type,Level,Replicate), col ="ID", sep="~", remove=FALSE, na.rm = FALSE)
DATA_ColNames <- colnames(DATA)
```

```{r}
#| label: sample average response 
#Calculate the average response based on the repeated injections
DATA_MEAN_INJ <- DATA %>% group_by(ID) %>% summarise_at(c("Std_conc", "Response", "Peak_asymmetry", "Rt_analyte"), list(~mean(.),~sd(.), ~rsd(., as.percentage = TRUE)))

DATA_MEAN_INJ <- separate(DATA_MEAN_INJ, col = ID, sep ="~", into = c("Type","Level", "Replicate"), remove = FALSE, )
```

## Validation Parameters

This section outlines the parameters tested, acceptance criteria, and results.

### Specificity

Procedure: Analyze blank, placebo, spiked samples, and degraded products.

Acceptance Criteria: No interference at the analyte's retention time.

Results: \[Provide chromatograms and qualitative observations.\]

### Linearity and Range

```         
My name is `r params$name`  \newcommand{\myname}{`r params$name`}   \newpage
```

Procedure: Prepare and analyze at least 5 concentrations.

Acceptance Criteria: Correlation coefficient (R² ≥ 0.99).

Results: \[Include calibration curve, equation, and R² value.\]

```{r}
#| label: Linear Regression
#Spilt the DATA tibble by StdReplicate #, into a list of tibbles
Tbl_Split <- DATA_MEAN_INJ %>%  
              filter(Type == "Standard") %>%  
              split(.$Replicate)

# fit each StdReplicate to a lm, makes a list
LM_STD_each <- Tbl_Split %>%
              map(~lm(Std_conc_mean ~ Response_mean, data = .)) 
GLM_STD_EACH <- LM_STD_each %>%
                map(~glance(.))

# make a lm of all replicates together
Tbl_STD_ALL <-  DATA_MEAN_INJ %>%  
                filter(Type == "Standard")

LM_STD_ALL <-Tbl_STD_ALL %>% 
  lm(Std_conc_mean ~ Response_mean, data = .) 

GLM_STD_ALL <- glance(LM_STD_ALL) 

              



```

```{r}
#| label: Calibration Curve Plots
#| 

#Plot the data and linear regression, color by Replicate
STD_CURVE_PLOT <- Tbl_STD_ALL %>% 
   ggplot( mapping = aes(x= Std_conc_mean, y = Response_mean)) +
  geom_smooth(method = "lm" ,se=T) 

STD_CURVE_PLOT +   
    geom_point(aes(color = Replicate)) +
  geom_smooth(aes(color =Replicate, fill = Replicate), method = "lm",se=F) +   
  labs(x=  paste("Standard Concentration of ", params$comp, " (", params$conc_units, ")", sep =""), y ="Response", title = "Calibration curve")
 

                                                                       
                                                                       
                                                                       
 


```

```{r}
#| Label: Cal Curve Statistics
library(dplyr) 
library(plyr)

# LM coeffiecents  for ALL STD as tibble
CO_STD_ALL <-LM_STD_ALL %>%
            coef(.)  %>%
            tidy(.) %>% 
            rename(., "ALL" = x, "Statistic" = "names")

# LM coeffiecents  for EACH STD as tibble
#Create MR function for extracting coefficients
getCoef <- function(df) {
          model <- lm(Std_conc_mean ~ Response_mean, data = df)
          coefs <- model$coef
          names(coefs) <- c("b", "m")
          R2 <- glimpse(model$r.squared)
          names(R2) <- c("R2")
          pvalue <-glimpse(model$sigma)
          names(sigma) <- c("Fstat")
          c(coefs, R2, Fstat)
          }
        
coefs.all <- getCoef(DATA_MEAN_INJ)

#MODEL1 <- tidy(LM_STD_each$`1`)
CO_STD_EACH <- LM_STD_each %>% 
                map(tidy(), .)  %>%
                rename(, "ALL" = x, "Statistic" = "names")

STATS_STD_ALL <- GLM_STD_ALL %>%
            select(.,"r.squared" , "statistic", "p.value", ) %>%             transpose_df(.) %>% rename(., "ALL" = "1", "Statistic" = "rowname" )
 Combined_ALL_STD  <- bind_rows(CO_STD_ALL, STATS_STD_ALL)  %>%                      mutate(., Statistic = c("b", "m", "R squared", "F-statistic", "p value"))
# now add individual Regressions to table 
# make the tibble into a table 

Combined_ALL_STD %>% 
  gt() %>%
  fmt_markdown(columns = everything()) %>%
fmt_number(columns ="ALL", n_sigfig = 5) %>% 
  fmt_scientific(columns ="ALL", n_sigfig = 5)

```

Calabration Curve Statistics

### Accuracy

Procedure: Spike analyte in matrix at 80%, 100%, and 120% of target concentration.

Acceptance Criteria: 98–102% recovery.

Results: \[Include % recovery data and statistical summary.\]

### Precision

### Repeatability:

Multiple replicates (n=6) of the same sample analyzed in one day.

### Intermediate Precision:

Tests on different days or with different analysts.

Acceptance Criteria: % RSD ≤ 2%.

Results: \[Provide % RSD for repeatability and intermediate precision.\]

### Limit of Detection (LOD) and Limit of Quantification (LOQ)

Procedure: Based on signal-to-noise ratio or calibration curve.

Acceptance Criteria: Clearly detectable and quantifiable concentrations.

esults: \[Provide calculated LOD and LOQ values.\]

### Robustness

Procedure: Vary method parameters (e.g., flow rate, temperature, pH) within small ranges.

Acceptance Criteria: No significant effect on results.

Results: \[Provide data showing robustness.\]

### System Suitability

Procedure: Assess retention time, theoretical plates, tailing factor, resolution.

Acceptance Criteria: \[Specify, e.g., Rs \> 2, Tailing Factor: 0.8–2.0\].

Results: \[Include system suitability test results.\]

### Stability of Analytical Solutions

Procedure: Test analyte stability over time under specified conditions.

Acceptance Criteria: Analyte remains within ±2% of initial value.

Results: \[Provide stability data.\]
